Disk Heating, Galactoseismology, and the Formation of Stellar Halos

### 2017-06-28 paper/paper.tex
last sections

% APW COMMENT: what exactly do we mean by "local"? < 2 kpc? < a few kpc?
Coincident with studies exploring structures at the very outer edge of our
Galactic disk, large scale spectroscopic surveys have allowed a detailed
re-examination of the {\it local} distribution of stellar velocities.
Using data from the SDSS \cite{widrow12,yanny13} and the RAdial Velocity
Experiment \cite[RAVE;][]{rave,williams13}, asymmetries between the northern and
southern Galactic hemispheres have been seen in the density and velocity
distributions of stars in the vicinity of the sun.
Looking $\sim$2 kpc out towards the Galactic anticenter, the Large Sky Area Multi-Object Fiber Spectroscopic Telescope \cite[LAMOST;][]{cui12,deng12,zhao12} finds similar asymmetries in radial and vertical velocities \cite{carlin13}.
The scale and sense of these asymmetries indicate moderate systematic motions (of order a few km/s) of stars within the disk perpendicular to the plane, suggesting both vertical movement of the midplane, and compression and rarefaction of the vertical scale \cite[referred to as ``bending'' and ``breathing'' modes respectively --- see, e.g.,][]{widrow14}.
It is natural to associate these asymmetries in motion as a local manifestations of the oscillations traced in space over much larger scales \cite{xu15,pricewhelan15}.
%{\it Summary: small-scale, systematic vertical motions of and within the disk have been detected in the Solar Neighborhood.}

@@

The observations summarized in Section \ref{sec:obs} indicate that:
\begin{itemize}
\item the low latitude structures --- Mon/GASS, TriAnd 1/2 and A13  --- each have low velocity dispersions supporting the genuine association of the stars within them;
\item Mon/GASS, TriAnd 1/2 and A13 also share a single, continuous sequence in average velocity as a function of $l$ suggestive of associations across these groups;
\item the stellar populations (as indicated by \frrmg) in these groups look like each other;
% ALLY COMMENT: Combine with next point? “look like each other” is a bit vague!
\item the stellar populations are more consistent with those in the Galactic disk rather than those observed in the stellar halo or Galactic  satellites;
\item the abundance patterns  of the low-latitude structures are similar to those of the thick disk of our Galaxy;
\item the low-latitude structures around the outer Galaxy can be connected to overdensities oscillating between the northern and southern hemispheres on smaller scales and traced all the way back to the Solar Neighborhood;
\item small-scale, systematic vertical motions of and within the disk have been detected in the Solar Neighborhood which suggest the local disk is moving up and down.
\end{itemize}

@@

% TING COMMENT: everything above line 251 is really summary of observation.
% Maybe this should go to Section 2.6? It is kinda confusing that this section
% is for theory but it were all about observations for the first 20 lines
Figure \ref{fig:cartoon} is a cartoon which summarizes all the observational works and their implications, by showing the approximate locations and amplitudes of the spatial and velocity structures that have been identified. (Note that the figure is misleading as the structures are {\it not} consistent morphologically with concentric rings that are axisymmetric about the Galactic center.)
Taken together, we conclude that: (i) there is compelling evidence that Mon/GASS, TriAnd 1/2 and A13 represent populations of stars formed in the disk that now exist at extreme $(R,Z)$;
(ii) these structures are associated with each other as part of a global system of vertical disk oscillations that can be traced all the way to the velocity asymmetries seen in the Solar Neighborhood; and
(iii) the stellar populations in and chemical abundances of these structures are inconsistent with a picture in which they formed from an accreted satellite.

One natural interpretation of these collected observations is that the oscillations represent the response of the disk to an external perturbation, for example as the impact of a satellite galaxy is transmitted and amplified by its wake in the dark matter halo \cite[as described for the LMC in][]{weinberg06}
Prior work has already pointed to this as a possible explanation for the existence of Mon/GASS \cite{kazantzidis08,younger08}, with the Sagittarius dwarf galaxy being pointed to as a plausible culprit for the perturbation \cite{purcell11}.
It has also been demonstrated how perturbations from a satellite on an orbit perpendicular to the disk could lead to bending (at low relative impact velocity) and breathing (at higher relative velocity) modes that would be observed in the Solar Neighborhood as asymmetries in the local velocity distribution  \cite{widrow14} and on larger scales as rings \cite{donghia16}.
Simulations have also shown that Sgr could be responsible for local velocity structure  \cite{gomez13}.
Such interactions and corresponding disk features have been found to naturally occur in cosmological simulations \cite{gomez16}.

\begin{figure}[ht!]
\centering
\includegraphics[width=5 in]{figures/simulation.pdf}
\caption{\label{fig:chervin}
}
\end{figure}

Figure \ref{fig:chervin} illustrates these ideas with the results of simulations from our own recent work.
We extended the prior theoretical backdrop that looked at Mon/GASS to examine whether the extreme locations of TriAnd 1/2 could fit within the same picture using simulations of a disk disturbed (separately) by satellites on orbits that mimic those expected for the Large Magellanic Cloud and Sagittarius \cite{laporte17}.
With different masses and orbits (and consequently different interaction  strength, timings, and durartions) these satellite necessarily induced distinct but overlapping signatures.
In  more recent work, we found that a model that was capable of reproducing the scales of the observed disturbances (radial wavelength and  amplitude in space, as well as magnitude of offsets in velocity locally) needed:
the  interaction of Sgr with the disk of the Milky Way to be followed for several passages longer than prior work;
Sgr needs to have sufficient initial mass and density to impact the disk in the last Gigayear with a remaining mass of $\sim3\times10^{9} M_{\odot}$;
and the disk to be realized with stars existing as far out as ???kpc form the Galactic center in order to populate the regions corresponding to TriAnd 1/2.
The interaction with the LMC modified the overall morphology of the structures induced, but was not sufficient alone to explain their properties.
The full details of these results will be discussed in an upcoming paper (Laporte et al, 2017).

@@

%It is interesting to place the current work in the context of ongoing and near-future studies of our Galaxy.
While the connections that have already been made between the different low-latitude structures and the disk population are convincing, there are several directions for further observations which would greatly facilitate an informative comparison to theoretical work, allowing more detailed interpretations and tighter constraints on matter distributions and histories.

The most obvious first direction would be to increase the dimensions of motion and accuracy of distance estimates to known features.
For example,  proper motions for the ``Anti-Center Stream'' (ACS) \cite[which may or may not be part of the larger Mon/GASS structure, see ] []{li12} indicate that stars in the ACS are not actually moving parallel to the stream \cite{carlin10}. This is inconsistent with expectations for the behavior of debris from a destroyed satellite.
If similar measurements of proper motions of stars in all the low-latitude structures showed significant motion perpendicular to the Galactic disk, this would provide the final evidence of a disk origin and connection to local oscillations, as well as important constraints on dynamical models (see Section \ref{sec:conc}).
In near-future data-releases, ESA's {\it Gaia} satellite \cite{???} is poised to provide just such proper motion data --- with proper motion accuracies of ??? $\mu$as/year or ?? km/s for our (closest) ??? mag M giant stars in Mon/GASS and ??? $\mu$as/year or ?? km/s for our (farthest) ??? mag M giant stars in TriAnd.

The second direction is to extend the various maps to to fainter magnitudes and global scales. Again, {\it Gaia} will be able to tackle this with distances, proper motions and radial velocities, although its reach towards low latitudes in the inner Galaxy will be limited by extinction.
Infrared surveys, such as APOGEE \cite{apogee}, could overcome this limitation and reach entirely across the Galaxy. Spectra from such a survey could provide both motions and abundance patterns.

% TING COMMENT: There are also a significant amount of studies on "galaxy
% morphology and warps" with distance galaxies. For example:
% http://adsabs.harvard.edu/abs/2016JKAS...49..239A
% You may want to mention these studies (reference therein) as well....
\subsection{Other Galaxies}

The great advantage of star-count studies is the ability to reach extremely low surface brightness levels. For example, the TriAnd clouds have been estimated to have surface brightness lower than
$\Sigma <$32 mag/arcsec$^2$ \cite{majewski04}.
%For example, TriAnd I contains ?? M-giant stars spread over an area on the sky of roughly ????, so ? giants/deg$^2$.
%Adopting a 10 Gyr-old isochrone for this population \cite[from the Padova group][]{}, each M-giant has an associated total stellar luminosity of ??? in the ?? band. Hence, the equivalanet urfc ebright ness would be .....

The growing samples of galaxies within and beyond the Local Group being mapped to extremely low surface brightness levels are intriguing.
Nearby, these can be reached, like the Milky Way, through star counts studies, most spectacularly for the case of our nearest neighbor, the Andromeda Galaxy, where giant star counts have revealed a significantly extended and richly featured outer stellar disk \cite{ferguson02,ibata05}.
Analogous studies have been carried out for galaxies up to distances of several Mpc \cite[e.g.,][]{monachesi13,crnojevi16}, although the focus of these studies has typically been on detecting the stellar halo of these objects.
Several dedicated surveys have made innovative advances in studying galaxies to low surface-brightness using a variety of techniques to reach limits below 30 mag/arcsec$^2$ in integrated light \citep[e.g.][]{delgado10,vandokkum14,duc15}.

@@

The above sections  summarized observational evidence for large scale vertical oscillations of the Galactic plane present in the Solar Neighborhood and reaching out beyond the traditional edge of our stellar disk.
They also collated theoretical studies that suggest that these oscillations could be caused by, and contain the signatures of, ongoing interactions of the Milky Way with its satellite system.
Moreover, observational prospects are bright for extending this work both to globally map the Milky Way, and to look for analogous features around many other galaxies.

Now that we have a physical picture of the origin of such features, as well as prospects for mapping them further within the Milky Way and detecting analogous substructures around other galaxies, we can move on to discussing how useful they are as probes of dynamics and history.
It is important to have these uses in mind to motivate and frame  future work.
While the mere existence of these structures is interesting, they contain a tiny fraction of the stars in galaxies spread out over a large area --- these properties naturally make them difficult to map, either because their unique signatures can be lost in the foreground star counts (e.g., in the Milky Way), or  because the required surface brightness limits for detection are prohibitively low (for integrated light).

Conversely, these features around the outskirts of galaxies may prove to be particularly powerful probes of interactions and histories, precisely because they contain so little mass --- they can be modeled rather simply as test particles responding to an external perturbation.

Below are just three examples of where these structures could promise new insights into some classic questions.
\begin{description}
\item{\it Disk heating mechanisms ---}
It has been understood for a long time that disks can evolve significantly due to mergers, major or minor, and hence that their current structures bear witness to their accretion history \cite{toth92,quinn93,walker96,velazquez99}.
This understanding has fueled a significant literature on the importance of the heating of galactic disks in response to encounters with other dark matter halos (that may or may not contain their own galaxies) \cite{font01,ardi03,benson04,stewart08,hopkins08,villalobos08,purcell09,kazantzidis09,sachdeva16,moetazedian16}.
In general, these studies have concentrated on the overall effects of many encounters on global properties, such as the thickness and vertical velocity dispersion in disks.
Their results have traditionally been compared to these scales in samples of galaxies.
In contrast, the identification and mapping of vertical waves in our Milky Way associated with ongoing interactions gives us the opportunity to dissect an individual disk heating event in progress.
We can use this to check our understanding of the mechanism directly and in detail rather than assessing its importance through collective effects and longterm, phase-mixed signatures.
\item{\it Stellar halo formation processes ----}
The last decade has seen increasing interest in assessing how much of the content of stellar halos could be made from stars originally formed in the disks of the galaxies that they surround. Hydrodynamical simulations of galaxy formaion suggest that tens of percent of the inner halo might be formed this way
\cite{abadi06,zolotov09,zolotov10,font11,mccarthy12,tissera13,tissera14,pillepich15,cooper15}.
Preliminary arguments for the existence of this ``kicked-out-disk'' population were based on transitions in the density or orbital structures of stellar halos \cite[e.g.,][]{carollo07}.
However, such transitions were also found to naturally occur in purely-accreted models of stellar halos \cite{deason13}.
More convincing observational evidence for disk stars in the halo is just beginning to emerge through studies with look for stars with halo-like kinematics, but disk-like abundances around M31 \cite{dorman13} and the Milky Way \cite{sheffield12,hawkins15,bonaca17}.
Our own work adds new perspectives on this stellar halo formation process with the detection and modeling of disk stars that may be in transition from the disk to the halo.
\item{\it Galactoseismic probes of interactions and dark matter ---}
The response of a disk to an encounter will depend on its own properties, the properties of the dark matter halo in which it is embedded and the mass and orbit of the perturbing satellite.
This leads to the suggestion that, analogous to helioseismic investigations of the structure of our Sun, maps of a disk response --- such as those described in Section \ref{sec:obs} for our Milky Way --- might be similarly inverted to tell us about (e.g.) the structure of the dark matter halo
\cite{widrow12}.
Indeed, recent studies of the signatures of encounters in the very outskirts of extended HI disks have successfully used simulations combined with an analytic understanding to find how the observed characteristics of the disturbed gas can be simply related to properties of the perturbing object \cite{chakrabarti09,chakrabarti11b,chang11}.
\end{description}

>>>>

% APW COMMENT: what exactly do we mean by "local"?
Coincident with studies exploring structures at the very outer edge of our
Galactic disk, large scale spectroscopic surveys have allowed a detailed
re-examination of the {\it local} distribution of stellar velocities.
Using data from the SDSS \cite{widrow12,yanny13} and the RAdial Velocity
Experiment \cite[RAVE;][]{rave,williams13}, asymmetries between the northern and
southern Galactic hemispheres have been seen in the density and velocity
distributions of stars in the vicinity of the sun.
Looking $\sim$2 kpc out towards the Galactic anticenter, the Large Sky Area Multi-Object Fiber Spectroscopic Telescope \cite[LAMOST;][]{cui12,deng12,zhao12} finds similar asymmetries in radial and vertical velocities \cite{carlin13}.
The scale and sense of these asymmetries indicate moderate systematic motions (of order a few km/s) of stars within the disk perpendicular to the plane, suggesting both vertical movement of the midplane, and compression and rarefaction of the vertical scale \cite[referred to as ``bending'' and ``breathing'' modes respectively --- see, e.g.,][]{widrow14}.
It is natural to associate these asymmetries in motion as a local manifestations of the oscillations traced in space over much larger scales \cite{xu15,pricewhelan15}.
%{\it Summary: small-scale, systematic vertical motions of and within the disk have been detected in the Solar Neighborhood.}

@@

% APW COMMENT: I think Chervin's simulation may be fine as our "cartoon" ... may
% be that the structures are too complex to simplify into a 1D cartoon

@@

% TING COMMENT: everything above line 251 is really summary of observation.
% Maybe this should go to Section 2.6? It is kinda confusing that this section
% is for theory but it were all about observations for the first 20 lines
% Figure \ref{fig:cartoon} is a cartoon which summarizes all the observational works and their implications, by showing the approximate locations and amplitudes of the spatial and velocity structures that have been identified. (Note that the figure is misleading as the structures are {\it not} consistent morphologically with concentric rings that are axisymmetric about the Galactic center.)

The observations summarized in Section \ref{sec:obs} indicate that:
\begin{itemize}
\item the low-latitude structures --- Mon/GASS, TriAnd, and A13  --- each have
      low velocity dispersions supporting the genuine association of the
      candidate member stars;
\item Mon/GASS, TriAnd, and A13 share a continuous sequence in mean GSR velocity
      as a function of Galactic longitude, suggestive of associations between
      these groups;
\item the stellar populations in the structures (as indicated by \frrmg) are all
      more consistent with those in the Galactic disk rather than those observed
      in the stellar halo or Galactic satellites;
\item the abundance patterns of stars in TriAnd and A13 are similar to those
      found in the thick disk of our Galaxy;
\item the low-latitude structures (around the outer disk) may be connected to
      oscillating density and velocity structure on smaller scales, traced all
      the way back to the solar neighborhood;
\end{itemize}
Taken together, we conclude that: (i) there is compelling evidence that
Mon/GASS, TriAnd, and A13 represent populations of stars formed in the disk that
now exist at extreme radii and heights above the Galactic disk;
(ii) these structures are associated and part of a global system of vertical
disk oscillations that can be traced all the way to the velocity asymmetries
seen in the solar neighborhood; and
(iii) the stellar populations in and chemical abundances of these structures are
inconsistent with a picture in which they formed from an accreted satellite.

One natural interpretation of these collected observations is that the oscillations represent the response of the disk to an external perturbation, for example as the impact of a satellite galaxy is transmitted and amplified by its wake in the dark matter halo \cite[as described for the LMC in][]{weinberg06}
Prior work has already pointed to this as a possible explanation for the existence of Mon/GASS \cite{kazantzidis08,younger08}, with the Sagittarius dwarf galaxy being pointed to as a plausible culprit for the perturbation \cite{purcell11}.
It has also been demonstrated how perturbations from a satellite on an orbit perpendicular to the disk could lead to bending (at low relative impact velocity) and breathing (at higher relative velocity) modes that would be observed in the solar neighborhood as asymmetries in the local velocity distribution  \cite{widrow14} and on larger scales as rings \cite{donghia16}.
Simulations have also shown that Sgr could be responsible for local velocity structure  \cite{gomez13}.
Such interactions and corresponding disk features have been found to naturally occur in cosmological simulations \cite{gomez16}.

\begin{figure}[ht!]
\centering
\includegraphics[width=5 in]{figures/simulation.pdf}
\caption{\label{fig:chervin}
% APW COMMENT: needs a caption - waiting until we have final version of figure.
}
\end{figure}

Figure \ref{fig:chervin} illustrates these ideas with the results of simulations from our own recent work.
Using simulations of a disk disturbed (separately) by satellites on orbits that
mimic those expected for the Large Magellanic Cloud and Sgr \cite{laporte17a}, we
extend the prior theoretical backdrop that looked at Mon/GASS to examine whether
the extreme locations of TriAnd stars could fit within the same picture.
With different masses and orbits (and consequently different interaction
strength, timings, and durations) these satellites necessarily induce distinct
but overlapping signatures on the global structure of the disk.
In more recent work, we found that a model that was capable of reproducing the scales of the observed disturbances (radial wavelength and amplitude in space, as well as magnitude of offsets in velocity locally) required:
the interaction of Sgr with the disk of the Milky Way to be followed for several passages longer than prior work;
Sgr must have sufficient initial mass and density to impact the disk in the last Gigayear with a remaining mass of $\sim3\times10^{9}~\msun$;
and the disk to be realized with stars existing as far out as $40~\kpc$ from the Galactic center in order to populate the regions corresponding to TriAnd.
The interaction with the LMC modified the overall morphology of the structures induced, but was not sufficient alone to explain their properties.
The full details of these results will be discussed in an upcoming paper (Laporte et al., in prep.).

@@

%It is interesting to place the current work in the context of ongoing and near-future studies of our Galaxy.
While the connections that have already been made between the different
low-latitude structures and the disk population are convincing, there are
several possible directions for further observations to strengthen these claims.
More detailed kinematic information and density measurements for more tracers
would greatly facilitate an informative comparison to theoretical work, allowing
more detailed interpretations of the history of our Galaxy.

The most obvious direction is to obtain proper motions and more accurate
distance estimates to the known features, using the candidate members discussed
in this work to search for other tracers.
For example, proper motions for the ``Anti-Center Stream'' (ACS) \cite[which may
or may not be part of the larger Mon/GASS structure;] []{li12} indicate that
stars in the ACS are not actually moving parallel to the stream \cite{carlin10}.
This is inconsistent with expectations for the behavior of debris from a
destroyed satellite.
If similar measurements of proper motions of stars in all of the low-latitude
structures showed significant motion perpendicular to the Galactic disk, this
would provide conclusive evidence of a disk origin and connection to local
oscillations.
With precise proper motions, the velocity information would also place important
constraints on dynamical models of the disk (see Section~\ref{sec:conc}).
In upcoming data-releases, astrometric data from the {\it Gaia} mission
\cite{gaia} is poised to provide these data.
Expected proper motion uncertainties for the closest M giant stars ($\approx
5~\kpc$) in Mon/GASS correspond to tangential velocity uncertainties of $\approx
1$--$2~\kms$ for M giants with tangential velocities $<50~\kms$.
For the farthest known M giant stars in TriAnd ($\approx 35~\kpc$), tangential
velocity uncertainties will be $\approx 7$--$12~\kms$ for M giants with
tangential velocities $<50~\kms$.

Another complimentary direction for future data is to extend the photometric and spectroscopic maps to fainter magnitudes and global scales.
{\it Gaia} will be able to tackle this with distances, proper motions, and
radial velocities, although its reach towards low latitudes in the inner Galaxy
will be limited by extinction.
Infrared surveys, such as APOGEE \cite{apogee}, could overcome this limitation
and reach entirely across the Galaxy. Spectra from such a survey could provide
both radial velocities and abundance patterns.

% APW COMMENT: you mention LSST in next section, but what about how LSST will
% provide deep photometry - map far down the main sequence for stars in these
% structures.

% APW COMMENT: I feel like we need to say more about APOGEE, future
% spectroscopic surveys like DESI, PFS.

\subsection{Other Galaxies}

% TING COMMENT: There are also a significant amount of studies on "galaxy
% morphology and warps" with distance galaxies. For example:
% http://adsabs.harvard.edu/abs/2016JKAS...49..239A
% You may want to mention these studies (reference therein) as well....

% APW COMMENT: this sentence comes a bit out of the blue -- needs a better
% introductory or transitional statement
The great advantage of star-count studies is the ability to reach extremely low
surface brightnesses. For example, TriAnd is estimated to have a surface
brightness lower than $\Sigma <$32 mag/arcsec$^2$ \cite{majewski04}.
%For example, TriAnd I contains ?? M-giant stars spread over an area on the sky of roughly ????, so ? giants/deg$^2$.
%Adopting a 10 Gyr-old isochrone for this population \cite[from the Padova group][]{}, each M-giant has an associated total stellar luminosity of ??? in the ?? band. Hence, the equivalanet urfc ebright ness would be .....

% APW COMMENT: this section needs some work, but ran out of steam...

The growing samples of galaxies within and beyond the Local Group being mapped to extremely low surface brightness levels are intriguing.
For nearby galaxies, these are through star counts studies, most spectacularly for the case of our nearest neighbor, the Andromeda Galaxy, where giant star counts have revealed a significantly extended and richly featured outer stellar disk \cite{ferguson02,ibata05}.
Analogous studies have been carried out for galaxies up to distances of several Mpc \cite[e.g.,][]{monachesi13,crnojevi16}, although the focus of these studies has typically been on detecting the stellar halo of these objects.
Several dedicated surveys have made innovative advances in studying galaxies to low surface-brightness using a variety of techniques to reach limits below 30 mag/arcsec$^2$ in integrated light \citep[e.g.][]{delgado10,vandokkum14,duc15}.

@@

The above sections summarize observational evidence for large scale vertical
oscillations of the Galactic plane present in the solar neighborhood and
reaching out beyond the traditional edge of our stellar disk.
We have also discussed theoretical studies that suggest that these oscillations
could be caused by, and contain the signatures of, ongoing interactions of the
Milky Way with its satellite system.
Moreover, there are numerous observational prospects for extending this work
both to globally map the Milky Way and to look for analogous features around
many other galaxies.

Now that we have a physical picture of the origin of such features, as well as prospects for mapping them further within the Milky Way and detecting analogous substructures around other galaxies, we can move on to discussing how useful they are for constraining the dynamics and evolution of galaxies.
While the mere existence of these structures is interesting, they contain a tiny fraction of the stars in galaxies spread out over a large area --- these properties naturally make them difficult to map, either because their unique signatures can be lost in the foreground star counts (e.g., in the Milky Way), or because the required surface brightness limits for detection are prohibitively low (for integrated light).

Conversely, these features around the outskirts of galaxies may prove to be particularly powerful probes of interactions and histories, precisely because they contain so little mass: they can be modeled as test particles responding to an external perturbation.

Below are just three examples of where these structures could promise new insights into some classic questions.
\begin{description}
\item{\it Disk heating mechanisms ---}
It has been understood for a long time that disks can evolve significantly due to mergers, major or minor, and hence that their current structures bear witness to their accretion history \cite{toth92,quinn93,walker96,velazquez99}.
This understanding has fueled a significant literature on the importance of the heating of galactic disks in response to encounters with other dark matter halos (that may or may not contain their own galaxies) \cite{font01,ardi03,benson04,stewart08,hopkins08,villalobos08,purcell09,kazantzidis09,sachdeva16,moetazedian16}.
In general, these studies have concentrated on the overall effects of many encounters on global properties, such as the thickness and vertical velocity dispersion in disks.
Their results have traditionally been compared to these scales in samples of galaxies.
In contrast, the identification and mapping of vertical waves associated with ongoing interactions in the Milky Way gives us the opportunity to dissect individual disk heating events in progress.
We can use this to check our understanding of the mechanism directly and in detail rather than assessing its importance through collective effects and longterm, phase-mixed signatures.
\item{\it Stellar halo formation processes ----}
The last decade has seen increasing interest in assessing how much of the content of stellar halos could be made from stars originally formed in the disks of the galaxies that they surround. Hydrodynamical simulations of galaxy formation suggest that tens of percent of the inner halo might be formed this way
\cite{abadi06,zolotov09,zolotov10,font11,mccarthy12,tissera13,tissera14,pillepich15,cooper15}.
Preliminary arguments for the existence of this ``kicked-out-disk'' population were based on transitions in the density or orbital structures of stellar halos \cite[e.g.,][]{carollo07}.
However, such transitions were also found to naturally occur in purely-accreted models of stellar halos \cite{deason13}.
More convincing observational evidence for disk stars in the halo is just beginning to emerge through studies that look for stars with halo-like kinematics, but disk-like abundances around M31 \cite{dorman13} and the Milky Way \cite{sheffield12,hawkins15,bonaca17}.
Our own work adds new perspectives on this stellar halo formation process with the detection and modeling of disk stars that may be in transition from the disk to the halo.
\item{\it Galactoseismic probes of interactions and dark matter ---}
The response of a disk to an encounter will depend on its own properties, the properties of the dark matter halo in which it is embedded and the mass and orbit of the perturbing satellite.
This leads to the suggestion that, analogous to helioseismic investigations of the structure of our Sun, maps of a disk response --- such as those described in Section \ref{sec:obs} for our Milky Way --- might be similarly inverted to tell us about (e.g.) the structure of the dark matter halo
\cite{widrow12}.
Indeed, recent studies of the signatures of encounters in the very outskirts of extended HI disks have successfully used simulations combined with an analytic understanding to find how the observed characteristics of the disturbed gas can be simply related to properties of the perturbing object \cite{chakrabarti09,chakrabarti11b,chang11}.
\end{description}
